Tropical mathematics and the Lambda-calculus I

### 2023-08-02 CSL24/sectionLaw.tex
commit

To relate this elegant correspondence with our work on $\LREL_{!}$, we show that it lifts to a model of the differential $\lambda$-calculus, extending the co-Kleisli category $\LREL_{!}$. 
First, we need to define a Lafont exponential $!$ over $\Mod$, and for this we use a well-known recipe from \cite{Mellies2018, Manzo2013}, that is: we first define a symmetric algebra $\Sym_{n}(M)$ as the equalizer of all permutative actions on $n$-tensors $M\otimes \dots \otimes M$; then, exploiting suitable properties of $\Mod$, we may define $!$ as the product of the symmetric algebras.  
This ensures that the coKleisli category $\Mod_{!}$ is a CCC.
Moreover, the constructions for $\Mod$ generalize those of $\LREL$, in the sense that $!(\Lawv^{X})\simeq \Lawv^{\multiset(X)}$ and that $\Mod_{!}(\Lawv^{X},\Lawv^{Y})\simeq \LREL_{!}(X,Y)$.
Finally, as the coKleisli category of a Lafont category with biproducts is always a CC$\partial$C \cite{LemayCALCO2021}, we can endow $\Mod_{!}$ with a suitable differential operator $E$ generalizing $D^{!}$ (see the Appendix) so that:
\begin{theorem}
($\Mod_{!}/\GMet_{!},E$) is a CC$\partial$C.
%, when equipped with 
%the differential operator $E$ defined, for $f:\ !M\to N$, by 
%$Ef(\alpha)=
%\bigvee\left\{
%f(\beta\cup [x]) \ \Big \vert  \ 
%\iota_{n}(\beta)\otimes \iota_{1}(x) \leq S(\alpha)
%\right\}
%$, where $\iota_{k}: M_{k}\to \prod_{i\in I}M_{i}$ is the injection morphism given by $\iota_{k}(x)( k)=x$ and $\iota_{k}(x)(i\neq k)=\infty$,
%and $S: \ !(M\times N)\to !M\otimes !N$ is the Seely isomorphism \cite{Mellies2018}.
\end{theorem}

>>>>

To relate this elegant correspondence with our work on $\LREL_{!}$, we show that it lifts to a model of the differential $\lambda$-calculus, extending the co-Kleisli category $\LREL_{!}$. 
First, we need to define a Lafont exponential $!$ over $\Mod$, and for this we use a well-known recipe from \cite{Mellies2018, Manzo2013}, that is: we first define a symmetric algebra $\Sym_{n}(M)$ as the equalizer of all permutative actions on $n$-tensors $M\otimes \dots \otimes M$; then, exploiting suitable properties of $\Mod$, we may define $!$ as the product of the symmetric algebras.  
This ensures that the coKleisli category $\Mod_{!}$ is a CCC.
Moreover, the constructions for $\Mod$ generalize those of $\LREL$, in the sense that $!(\Lawv^{X})\simeq \Lawv^{\multiset(X)}$ and that $\Mod_{!}(\Lawv^{X},\Lawv^{Y})\simeq \LREL_{!}(X,Y)$.
Finally, as the coKleisli category of a Lafont category with biproducts is always a CC$\partial$C \cite[Theorem 21]{LemayCALCO2021}, we can endow $\Mod_{!}$ with a  differential operator $E$ generalizing $D^{!}$ (see the Appendix) so that:
\begin{theorem}\label{thm:lemay}
($\Mod_{!}/\GMet_{!},E$) is a CC$\partial$C.
%, when equipped with 
%the differential operator $E$ defined, for $f:\ !M\to N$, by 
%$Ef(\alpha)=
%\bigvee\left\{
%f(\beta\cup [x]) \ \Big \vert  \ 
%\iota_{n}(\beta)\otimes \iota_{1}(x) \leq S(\alpha)
%\right\}
%$, where $\iota_{k}: M_{k}\to \prod_{i\in I}M_{i}$ is the injection morphism given by $\iota_{k}(x)( k)=x$ and $\iota_{k}(x)(i\neq k)=\infty$,
%and $S: \ !(M\times N)\to !M\otimes !N$ is the Seely isomorphism \cite{Mellies2018}.
\end{theorem}
